Numerical simulations of the Cordilleran ice sheet through the last glacial cycle

### 2016-01-22 cordillera-cycle.tex
Rev. Jarosch: clarify ETOPO1 bilinear interp.
Rev. Jarosch: address limits of ETOPO1 data.

      Basal topography is derived from the ETOPO1 combined topography and
      bathymetry dataset with a~resolution of 1\,arc-min
      \citep{Amante.Eakins.2009}. Sea level is lowered as a~function of time
      based on the Spectral Mapping Project
      \citep[SPECMAP,][]{Imbrie.etal.1989} time scale. Geothermal heat flux
      is applied as a~constant value of
      $q_{\mathrm{G}}=70$\,\unit{mW\,m^{-2}} at
      3\,\unit{km} depth (Sect.~\ref{sec:icedyn}). Basal sliding follows
      a pseudo-plastic law where the yield stress accounts for till
      deconsolidation under high water pressure (Sect.~\ref{sec:sliding}).
      Ice shelf calving is calculated based on ice thickness and principle
      strain rates (Sect.~\ref{sec:sliding}). Surface mass balance is
      computed using a~positive degree-day (PDD) model
      (Sect.~\ref{sec:surface}). Default parameter values are given in
      Table~\ref{tab:params}. Climate forcing is provided by a~monthly
      climatology averaged from 1979 to 2000 from the North American
      Regional Reanalysis \citep[NARR,][]{Mesinger.etal.2006}, perturbated
      by time-dependent offsets and lapse-rate temperature corrections
      (Sect.~\ref{sec:atm}, Table~\ref{tab:records}). A sensitivity study
      to some of the most influential ice rheology (Sect~\ref{sec:icedyn})
      and basal sliding (Sect.~\ref{sec:sliding}) parameters was performed
      (Table~\ref{tab:sens_params}).

@@

      It is generally believed that the Cordilleran ice sheet formed by the
      coalescence of several mountain-centred ice caps
      \citep{Davis.Mathews.1944}. In our simulations, major ice-dispersal
      centres, visible on the modelled ice cover duration maps
      (Fig.~\ref{fig:hr_geom_duration}), are located over the Coast
      Mountains (CM), the Columbia and Rocky mountains (CRM), the Skeena
      Mountains (SM), and the Selwyn and Mackenzie mountains (SMKM). The
      Wrangell and Saint Elias mountains, heavily glacierized at present,
      host an ice cap for the entire length of both simulations, but that
      ice cap does not appear to be a~major feed to the Cordilleran ice
      sheet (Fig.~\ref{fig:hr_geom_duration}, WSEM). Although the Coast,
      Skeena and Columbia and Rocky mountains (CM, SM, CRM) are covered by
      mountain glaciers for most of the last glacial cycle, providing
      durable nucleation centres for an ice sheet, this is not the case for
      the Selwyn and Mackenzie mountains (SMKM), where ice cover on the
      highest peaks is limited to a~small fraction of the last glacial
      cycle. In other words, the Selwyn and Mackenzie mountains only appear
      as a~secondary ice-dispersal centre during the coldest periods of the
      last glacial cycle. The Northern Rocky Mountains
      (Fig.~\ref{fig:hr_geom_duration}, NRM) do not act as a~nucleation
      centre, but rather as a~pinning point for the Cordilleran ice sheet
      margin coming from the west.

>>>>

      Basal topography is bilinearly interpolated from the ETOPO1 combined
      topography and bathymetry dataset with a~resolution of 1\,arc-min
      \citep{Amante.Eakins.2009} to the model grids. Sea level is lowered as
      a~function of time based on the Spectral Mapping Project
      \citep[SPECMAP,][]{Imbrie.etal.1989} time scale. Geothermal heat flux
      is applied as a~constant value of
      $q_{\mathrm{G}}=70$\,\unit{mW\,m^{-2}} at
      3\,\unit{km} depth (Sect.~\ref{sec:icedyn}). Basal sliding follows
      a pseudo-plastic law where the yield stress accounts for till
      deconsolidation under high water pressure (Sect.~\ref{sec:sliding}).
      Ice shelf calving is calculated based on ice thickness and principle
      strain rates (Sect.~\ref{sec:sliding}). Surface mass balance is
      computed using a~positive degree-day (PDD) model
      (Sect.~\ref{sec:surface}). Default parameter values are given in
      Table~\ref{tab:params}. Climate forcing is provided by a~monthly
      climatology averaged from 1979 to 2000 from the North American
      Regional Reanalysis \citep[NARR,][]{Mesinger.etal.2006}, perturbated
      by time-dependent offsets and lapse-rate temperature corrections
      (Sect.~\ref{sec:atm}, Table~\ref{tab:records}). A sensitivity study
      to some of the most influential ice rheology (Sect~\ref{sec:icedyn})
      and basal sliding (Sect.~\ref{sec:sliding}) parameters was performed
      (Table~\ref{tab:sens_params}).

@@

      It is generally believed that the Cordilleran ice sheet formed by the
      coalescence of several mountain-centred ice caps
      \citep{Davis.Mathews.1944}. In our simulations, major ice-dispersal
      centres, visible on the modelled ice cover duration maps
      (Fig.~\ref{fig:hr_geom_duration}), are located over the Coast
      Mountains (CM), the Columbia and Rocky mountains (CRM), the Skeena
      Mountains (SM), and the Selwyn and Mackenzie mountains (SMKM).

      The location of modelled ice-dispersal centres is potentially biased
      by present day ice volumes contained in the ETOPO1 basal topography
      data. Although a reconstruction exist for the thickness of western
      Canadian glaciers south of 60{\degree}N \citep{Clarke.etal.2013}, the
      most problematic part of the model domain in this respect is by far
      that of the Wrangell and Saint Elias mountains, where ice thicknesses
      up to 1200\,m have been measured by low-frequency radar
      \citep{Rignot.etal.2013}. In this area, located over the USA Canada
      border just north of 60{\degree}N, temperate ice, surge dynamics and
      deep subglacial depressions in the icefield interior pose a
      fundamental challenge to reconstructing basal topography for the
      entire ice cap \citep{Rignot.etal.2013}. Although this causes our
      simulations to overestimate ice surface topography in this region, we
      are not aware of bed topography reconstructions that could be used to
      force the ice sheet model. However, this drawback seem to have
      little effect on modelled Cordilleran ice sheet dynamics. In fact, the
      Wrangell and Saint Elias mountains, heavily glacierized at present,
      host an ice cap for the entire length of both simulations, but that
      ice cap does not appear to be a~major feed to the Cordilleran ice
      sheet (Fig.~\ref{fig:hr_geom_duration}, WSEM). Moreover, with this
      exception of the Wrangell and Saint Elias mountains ice field,
      present-day ice volumes are small relative to the ice volumes
      concerned in our study.

      Although the Coast,
      Skeena and Columbia and Rocky mountains (CM, SM, CRM) are covered by
      mountain glaciers for most of the last glacial cycle, providing
      durable nucleation centres for an ice sheet, this is not the case for
      the Selwyn and Mackenzie mountains (SMKM), where ice cover on the
      highest peaks is limited to a~small fraction of the last glacial
      cycle. In other words, the Selwyn and Mackenzie mountains only appear
      as a~secondary ice-dispersal centre during the coldest periods of the
      last glacial cycle. The Northern Rocky Mountains
      (Fig.~\ref{fig:hr_geom_duration}, NRM) do not act as a~nucleation
      centre, but rather as a~pinning point for the Cordilleran ice sheet
      margin coming from the west.
